A Toolbox for Manipulating and Assessing Color Palettes

### 2021-04-08 pkg/colorspace/paper/Toolbox/article.Rmd
all URLs for hclwizard.org and colorspace.R-Forge.R-project.org now with https instead of http
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/colorspace@588 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

---
author:
  - name: Achim Zeileis
    # primary affiliation
    affiliation: Universität Innsbruck
    # full address
    address: |
      | Department of Statistics
      | Faculty of Economics and Statistics
      | Universitätsstr. 15
      | 6020 Innsbruck, Austria
    email: \email{Achim.Zeileis@R-project.org}
    url: https://eeecon.uibk.ac.at/~zeileis/
  - name: Jason C. Fisher
    affiliation: U.S. Geological Survey
  - name: Kurt Hornik
    affiliation: |
      | WU Wirtschafts-
      | universität Wien
  - name: Ross Ihaka
    affiliation: University of Auckland \AND
  - name: Claire D. McWhite
    affiliation: |
      | The University of
      | Texas at Austin
  - name: Paul Murrell
    affiliation: University of Auckland
  - name: Reto Stauffer
    affiliation: Universität Innsbruck
  - name: Claus O. Wilke
    affiliation: |
      | The University of
      | Texas at Austin
title:
  # title in title case
  formatted: "\\pkg{colorspace}: A Toolbox for Manipulating and Assessing Colors and Palettes"
  # without LaTeX markup (if any)
  plain:     "colorspace: A Toolbox for Manipulating and Assessing Colors and Palettes"
  # with LaTeX markup (if any), used as running title
  short:     "\\pkg{colorspace}: Manipulating and Assessing Colors and Palettes"
abstract: >
  The \proglang{R} package \pkg{colorspace} provides a
  flexible toolbox for selecting individual colors or color palettes, manipulating
  these colors, and employing them in statistical graphics and data
  visualizations. In particular, the package provides a broad range of color
  palettes based on the HCL (Hue-Chroma-Luminance) color space. The three HCL
  dimensions have been shown to match those of the human visual system very well,
  thus facilitating intuitive selection of color palettes through trajectories in
  this space.

@@

The graphics resulting from the demos can also be viewed online at
<http://colorspace.R-Forge.R-project.org/articles/approximations.html>.

@@

Different kinds of limitations can be emulated using the physiologically-based model
for simulating color vision deficiency (CVD) of @color:Machado+Oliveira+Fernandes:2009:
deuteranomaly (green cone cells defective),
protanomaly (red cone cells defective), and
tritanomaly (blue cone cells defective).
While most other CVD simulations handle only dichromacy, where one of three cones is
non-functional, @color:Machado+Oliveira+Fernandes:2009 provide a unified model of
both dichromacy and anomalous trichromacy, where one cone has shifted spectral
sensitivity. As anomalous trichromacy is the most common form of color vision
deficiency, it is important to emulate along with the rarer, but more severe
dichromacy. Below we briefly describe our \proglang{R} interface to these emulation
techniques and show them in practice for a heatmap with sequential palette.
Another example with a diverging palette is available at
<http://colorspace.R-Forge.R-project.org/articles/color_vision_deficiency.html>.
Finally, CVD emulation is particularly useful for bringing out why the RGB
rainbow palette is almost always a bad choice in scientific displays. See
<http://colorspace.R-Forge.R-project.org/articles/endrainbow.html> for further
illustrations.

@@

To facilitate exploring the package and employing it when working with colors,
several graphical user interfaces (GUIs) are provided within the package as
\pkg{shiny} apps [@color:shiny]. All of these GUIs/apps can be run locally
from within \proglang{R} and are also provided online at <http://hclwizard.org/>.

>>>>

---
author:
  - name: Achim Zeileis
    # primary affiliation
    affiliation: Universität Innsbruck
    # full address
    address: |
      | Department of Statistics
      | Faculty of Economics and Statistics
      | Universitätsstr. 15
      | 6020 Innsbruck, Austria
    email: \email{Achim.Zeileis@R-project.org}
    url: https://www.zeileis.org/
  - name: Jason C. Fisher
    affiliation: U.S. Geological Survey
  - name: Kurt Hornik
    affiliation: |
      | WU Wirtschafts-
      | universität Wien
  - name: Ross Ihaka
    affiliation: University of Auckland \AND
  - name: Claire D. McWhite
    affiliation: |
      | The University of
      | Texas at Austin
  - name: Paul Murrell
    affiliation: University of Auckland
  - name: Reto Stauffer
    affiliation: Universität Innsbruck
  - name: Claus O. Wilke
    affiliation: |
      | The University of
      | Texas at Austin
title:
  # title in title case
  formatted: "\\pkg{colorspace}: A Toolbox for Manipulating and Assessing Colors and Palettes"
  # without LaTeX markup (if any)
  plain:     "colorspace: A Toolbox for Manipulating and Assessing Colors and Palettes"
  # with LaTeX markup (if any), used as running title
  short:     "\\pkg{colorspace}: Manipulating and Assessing Colors and Palettes"
abstract: >
  The \proglang{R} package \pkg{colorspace} provides a
  flexible toolbox for selecting individual colors or color palettes, manipulating
  these colors, and employing them in statistical graphics and data
  visualizations. In particular, the package provides a broad range of color
  palettes based on the HCL (Hue-Chroma-Luminance) color space. The three HCL
  dimensions have been shown to match those of the human visual system very well,
  thus facilitating intuitive selection of color palettes through trajectories in
  this space.

@@

The graphics resulting from the demos can also be viewed online at
<https://colorspace.R-Forge.R-project.org/articles/approximations.html>.

@@

Different kinds of limitations can be emulated using the physiologically-based model
for simulating color vision deficiency (CVD) of @color:Machado+Oliveira+Fernandes:2009:
deuteranomaly (green cone cells defective),
protanomaly (red cone cells defective), and
tritanomaly (blue cone cells defective).
While most other CVD simulations handle only dichromacy, where one of three cones is
non-functional, @color:Machado+Oliveira+Fernandes:2009 provide a unified model of
both dichromacy and anomalous trichromacy, where one cone has shifted spectral
sensitivity. As anomalous trichromacy is the most common form of color vision
deficiency, it is important to emulate along with the rarer, but more severe
dichromacy. Below we briefly describe our \proglang{R} interface to these emulation
techniques and show them in practice for a heatmap with sequential palette.
Another example with a diverging palette is available at
<https://colorspace.R-Forge.R-project.org/articles/color_vision_deficiency.html>.
Finally, CVD emulation is particularly useful for bringing out why the RGB
rainbow palette is almost always a bad choice in scientific displays. See
<https://colorspace.R-Forge.R-project.org/articles/endrainbow.html> for further
illustrations.

@@

To facilitate exploring the package and employing it when working with colors,
several graphical user interfaces (GUIs) are provided within the package as
\pkg{shiny} apps [@color:shiny]. All of these GUIs/apps can be run locally
from within \proglang{R} and are also provided online at <https://hclwizard.org/>.
